---
title: "Kira Plastinina Customer Insights Project"
author: "Kennedy Muriuki"
date: "11/09/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Defining the question

In this weeks problem, I will be working as a data scientist for a client, Kira Plastinina. Kira Plastinina is a Russian brand that is sold throughout retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. It's marketing team would like to better understand their customers behaviors and therefore requested me to draw insights on the characteristics of various customer groups.


## Defining the metric of success

The metric of success is to obtain several distinct clusters of the Kira Plastina's customers and their characteristics.

## Data Sourcing

The dataset was sourced from the company's database of existing customers. The dataset was availed by the brand's sales and marketing team.

## The Experimental Design

For me to be able to obtain the required results, the following steps will be undertaken

1. Problem definition
2. Data sourcing
3. Checking the data
4. Performing data cleaning
5. Perform Exploratory data analysis
6. Implementing the solution
7. Challenging the solution
8. Follow up questions


## Checking the data

### Loading the dataset

```{r}
# the dataset has been downloaded to a local repository and will be loaded as a csv file
data <- read.csv(file.choose())
head(data)
```

```{r}
# previewing the bottom of the dataset
tail(data)
```


```{r}
# displaying the structure of the dataset
str(data)
```


```{r}
# displaying the dimension of the dataset
dim(data)
```
the data contains 12,330 entries and 18 columns

```{r}
# checking the names of the columns and their datatypes in the dataset

columns = colnames(data)
for (column in seq(length(colnames(data)))){
  print(columns[column])
  print(class(data[, column]))
  cat('\n')
}
```


### Checking for missing values

```{r}
# checking if the dataset contains any missing values
any(is.na(data))
```

```{r}
# checking the columns with missing data
colSums(is.na(data))
```


```{r}
# since the missing data is not a lot I will skip the missing data
df <- na.omit(data)

# cheking the dimension of the new dataset
dim(df)
```


### checking for duplicates

```{r}
# checking for duplicated data in the dataset
any(duplicated(df))
```


```{r}
# identifying the duplicated data
dup <- df[duplicated(df),]
dup
```

data showed above as duplicated did not look like duplicated data. They had a lot of similar entries in some columns but did not have entirely similar column entries. Therefore I will not remove the 177 rows as this might cause inconsistencies within the data set and affect final results

### checking for outliers

```{r }
# obtaining the numerical columns
numerical = df[,c(1:10)]
head(numerical)
```

```{r }
# generating boxplots for the numerical columns
par(mfrow=c(2,2), mar=c(5,4,2,2))

for (i in names(numerical)){
  x <- (numerical)[,i]
  boxplot(x, xlab= i, col="red")
  boxplot.stats(x)$out
}

```


```{r}
# dealing with the outliers
```


```{r}
# checking the summary of the dataset
summary(df)
```


## Univariate Exploratory Data Analysis

### Measures of central tendancies

```{r }
# obtaining the mean of the numerical columns

for (i in names(numerical)){
  x <- numerical[,i]
  mean <- mean(x)
  print(paste("The mean ", i , "is" ,  mean))
  cat('\n')
}
```


```{r }
# obtaining the median of the numerical columns

for (i in names(numerical)){
  x <- numerical[,i]
  mean <- median(x)
  print(paste("The mean ", i , "is" ,  mean))
  cat('\n')
}
```


```{r }
# displaying the mode of the numerical columns

getmode <- function(a){
  uniqv <- unique(a)
  uniqv[which.max(tabulate(match(a,uniqv)))]
}

# looping through the columns to get the mode

for (i in names(numerical)){
  x <- numerical[,i]
  mode <- getmode(x)
  print(paste("column", i , ":" ,  mode))
  cat('\n')
}
```


### Measures of dispersion

```{r }
# obtaining the five number summary of the numerical columns

for (i in names(numerical)){
  x <- numerical[,i]
  quantile <- quantile(x)
  print(paste(i))
  cat('\n')
  print(quantile)
  cat('\n')
}

```


```{r}
# showing the variances and standard deviation of the munerical columns
for (i in names(numerical)){
  x <- numerical[,i]
  Sdev <- sd(x)
  var <- var(x)
  print(paste(i))
  cat('\n')
  print(paste("Variance :", round(var, digits = 2), "Standard deviation :", round(Sdev, digits = 2)))
  cat('\n')
}


```
### Histograms


```{r}
# creating a function to generate histogram plots for the numerical columns
library(ggplot2)
dens_fun = function(X) {
     ggplot(numerical, aes(x= X)) +
      geom_histogram(color="blue", alpha=0.8)
}

# generating histogram for Administrative
dens_fun(df$Administrative)
```


```{r}
dens_fun(numerical$Administrative_Duration)
```


```{r}
dens_fun(numerical$Informational)
```


```{r}
dens_fun(numerical$Informational_Duration)
```


```{r}
dens_fun(numerical$ProductRelated )
```


```{r}
dens_fun(numerical$ProductRelated_Duration)
```


```{r}
dens_fun(numerical$BounceRates)
```


```{r}
dens_fun(numerical$ExitRates)
```


```{r}
dens_fun(numerical$PageValues)
```


```{r}
#dens_fun(numerical$SpecialDay)
ggplot(numerical, aes(x= SpecialDay)) +
  geom_histogram(binwidth= 0.15, color="blue")
```


### Bar graphs

```{r}
head(df)
```

```{r}
# visualizing the column months

ggplot(df, aes(Month)) + 
  geom_bar(fill = "green")
```

```{r}
# visualizing column OperatingSystems

ggplot(df, aes(OperatingSystems)) + 
  geom_bar(fill = "red")

```


```{r}
# visualizing browser column
ggplot(df, aes(Browser)) + 
  geom_bar(stat="count", fill = "green")
```


```{r}
# visualizing column OperatingSystems

ggplot(df, aes(Region)) + 
  geom_bar(fill = "blue")

```

```{r}
# visualizing column OperatingSystems

ggplot(df, aes(TrafficType)) + 
  geom_bar(fill = "grey")
```


```{r}
# plotting the graph of visitor type

ggplot(df, aes(VisitorType)) + 
  geom_bar(stat="count", fill = "green")
```

```{r}
# visualizing the weekend column
ggplot(df, aes(Weekend)) + 
  geom_bar(stat="count", fill = "black")
```


```{r}
# visualizing the revenue column
ggplot(df, aes(Revenue)) + 
  geom_bar(stat="count", fill = "orange")
```



## Bivariate analysis

```{r}
# performing a correlation plot for the numerical variables

library(ggcorrplot)

corr <- round(cor(numerical), 1)
head(corr,4)

ggcorrplot(corr, hc.order = TRUE, outline.col = "white")
```

```{r}
# perfoming pairplots for the numerical variables
col <- df[,c(1:10)]
pairs(col)
```



## Implimenting the solution

### K-Means Clustering

This method involves partitioning the data set into clusters or k groups.
```{r}
# Since clustering is unsupervised learning I will remove the class attribute which is revenue
df_new <- df[,c(1:9)]
df_class <- df[,c(10)]
```

```{r}
# looking at the predictor data
head(df_new)
```

```{r}
# normalizing the dataset
head(df)
```


```{r}
# printing unique values for the categorical columns
for (i in names(df[,c(11:18)])){
  print(unique(df[i]))
}
```


One hot encoding the categorical columns

```{r}
library(caret)
dummy <- dummyVars("~.", data=df, fullRank=T)
df_enc <- data.frame(predict(dummy, newdata=df))
str(df_enc)
```


```{r}
# scaling the numerical columns
df_scale <- scale(df_enc[,1:10])
head(df_scale)
```


```{r}
# joining the two dataset 
final <- cbind(df_scale, df_enc[,11:27])
head(final)
```

```{r}
# removing the class attribute
final_att <- final[,c(1:26)]
final_class <- final$RevenueTRUE
head(final_att)
```


```{r}
# applying the k-means clustering
result <- kmeans(final_att, 2)
result$size
result$centers

```


```{r}
# showing the table of the clusters
table(result$cluster, final_class)
```


```{r}
# visualizing the clusters
# installing the package
# install.packages("factoextra")
png("C:\\plot1.png", width = 480, height = 480, units = "px", bg = "white")
library(factoextra)

fviz_cluster(result, data = final_att)
```



```{r}
# obtaining the optimal number of clusters
fviz_nbclust(x = final_att, FUNcluster = kmeans, method = 'silhouette' )
```

### Hierachical clustering 

```{r}
# computing the eucledian distance between observations

d <- dist(final_att, method = "euclidean")
```


```{r}
# performing hierachical clustering usig hclust
result_hc <- hclust(d, method = "ward.D2")

```


```{r}
# plotting the dendograms
plot(result_hc, cev=0.6)
```

